---
title: "kNN_part_0_cleaning_up_origin_file"
author: "Sven Rodne"
date: "2025-12-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clear the environment if needed

```{r}
rm(list=ls())
```

Load in the required packages

```{r}
library(dplyr)
library(tidyverse)
library(mgcv)
library(ggplot2)
library(here)
```

Clean up the origin file and give full rows to all the tallied trees

```{r}
# Pulling original data copy to copy the new rows of trees based on tally values from the tally, x0_5_cm (meaning number of trees that are between 0 and 5 cm), or x5_10_cm columns (trees between 5 and 10 cm trees)
original_data <- read.csv("data_origin/rogue_basin_data_11_21_2023.csv")

# Convert NA tally values to 1 (assuming that this would mean they were trees that were single stemmed)
original_data <- original_data %>%
  mutate(tally = ifelse(is.na(tally), 1, tally))
```

Here I filter out trees without diam_cm (DBH) but have dstmp (diameter at stump height) to calculate their DBH values from a regression for trees with both diam_cm and dstmp

```{r}
# Filtering out the records without diam_cm but with dstmp
dstmp <- original_data %>%
  filter(is.na(diam_cm) & dstmp>0)

# Take the inverse of the dstmp dataframe for joining after calculating the diam_cm from dstmp
inverse_dstmp <- anti_join(original_data, dstmp, by = "rodne_uid")

# Generalized Additive Model for the regression equation to calculate diam_cm from dstmp
gammod1 <- gam(diam_cm ~ dstmp, data = inverse_dstmp)
summary(gammod1)$r.sq
summary(gammod1) # The variables calculated from this GAM are used for calculating diam_cm for trees with only dstmp

# Calculate diam_cm for the trees with only dstmp
dstmp <- dstmp %>%
  mutate(diam_cm = (dstmp*0.906580)-2.149607)

# Join dstmp and inverse_dstmp back into the original_data
original_data <- bind_rows(dstmp, inverse_dstmp)
```

Add the coordinate information for all trees by first splitting the data into either stem and density (the coordinate data is stored for each tree within the stem maps and each plot center for density plots)

```{r}
rogue_density_data <- original_data %>%
  filter(plot_type=="density")

rogue_stem_data <- original_data %>%
  filter(plot_type=="stem")

# Stem spatial data
stem_spatial <- read.csv("csv_input_deck/stem_map_spatial_data.csv")
stem_spatial <- stem_spatial[,c("uid", "X", "Y")]
colnames(stem_spatial) <- c("uid", "x", "y")
rogue_stem_data <- merge(rogue_stem_data, stem_spatial, by = "uid", all.x = TRUE)
subsetroguestem <- rogue_stem_data[,c("uid", "x", "y")]

# Density spatial data
density_spatial <- read.csv("csv_input_deck/densityplotcoords.csv")
density_spatial <- density_spatial[,c("Rodne_PID", "POINT_X", "POINT_Y")]
colnames(density_spatial) <- c("rodne_pid", "x", "y")
density_spatial <- unique(density_spatial)
rogue_density_data <- merge(rogue_density_data, density_spatial, by = "rodne_pid", all.x = TRUE)
subsetroguedensity <- rogue_density_data[,c("iid", "x", "y")]

# Bind all spatial data together
all_rogue <- bind_rows(rogue_stem_data, rogue_density_data)

# Removing both the Star Gulch stem map and the Boaz Gulch site since this data is not used for the rest of the project
all_rogue <- all_rogue %>%
  filter(!site_name == "Boaz Gulch" & !rodne_pid == "SA")

# Revert the new dataframe back to "original_df" for continuity
original_df <- all_rogue
```

Now to expand all the times that tree tallies were used to create all the rows for each tree

```{r, echo=FALSE}
################################################################################

# For tally and diam_cm only

# Initialize an empty list to store duplicated rows
duplicated_rows <- list()

# Loop through each row of the original dataframe
for (i in 1:nrow(original_df)) {
  row <- original_df[i, ]
  
  # If tally has a number greater than 1, duplicate the row based on the value in tally
  if (!is.na(row$tally) && row$tally > 1) {
    for (j in 1:row$tally) {
      # Duplicate the row
      duplicated_row <- row
      
      # Increment uid and rodne_uid from "001" onwards
      duplicated_row$uid <- paste0(duplicated_row$uid, "_", sprintf("%03d", j))
      duplicated_row$rodne_uid <- paste0(duplicated_row$rodne_uid, "_", sprintf("%03d", j))
      
      # Append the duplicated row to the list
      duplicated_rows[[length(duplicated_rows) + 1]] <- duplicated_row
    }
  }
}

# Combine the duplicated rows into a dataframe
tallied_df <- do.call(rbind, duplicated_rows)

# Reset row names
rownames(tallied_df) <- NULL

# This results in the tallied_df

################################################################################

# For X0_5_cm

# Initialize an empty list to store duplicated rows
duplicated_rows <- list()

# Loop through each row of the original dataframe

for (i in 1:nrow(original_df)) {
  row <- original_df[i, ]
  
  # Print the value of X0_5_cm for debugging
  print(row$X0_5_cm)
  
  # If X0_5_cm has a number greater than 1, duplicate the row based on the value in X0_5_cm
  if (!is.na(row$X0_5_cm) && row$X0_5_cm > 1) {
    for (j in 1:row$X0_5_cm) {
      # Duplicate the row
      duplicated_row <- row
      
      # Increment uid and rodne_uid from "001" onwards
      duplicated_row$uid <- paste0(duplicated_row$uid, "_", sprintf("%03d", j))
      duplicated_row$rodne_uid <- paste0(duplicated_row$rodne_uid, "_", sprintf("%03d", j))
      
      # Set diam_cm to 2.5 for the duplicated row
      duplicated_row$diam_cm <- 2.5
      
      # Append the duplicated row to the list
      duplicated_rows[[length(duplicated_rows) + 1]] <- duplicated_row
    }
  }
  
  # If X0_5_cm has a value of 1, set diam_cm to 2.5
  if (!is.na(row$X0_5_cm) && row$X0_5_cm == 1) {
    duplicated_row <- row
    duplicated_row$diam_cm <- 2.5
    duplicated_rows[[length(duplicated_rows) + 1]] <- duplicated_row
  }
}

# Combine the duplicated rows into a dataframe
X0_5_cm_df <- do.call(rbind, duplicated_rows)

# Reset row names
rownames(X0_5_cm_df) <- NULL

################################################################################

# For X5_10_cm

# Initialize an empty list to store duplicated rows
duplicated_rows <- list()

# Loop through each row of the original dataframe
for (i in 1:nrow(original_df)) {
  row <- original_df[i, ]
  
  # If X5_10_cm has a number greater than 1, duplicate the row based on the value in X5_10_cm
  if (!is.na(row$X5_10_cm) && row$X5_10_cm > 1) {
    for (j in 1:row$X5_10_cm) {
      # Duplicate the row
      duplicated_row <- row
      
      # Increment uid and rodne_uid from "001" onwards
      duplicated_row$uid <- paste0(duplicated_row$uid, "_", sprintf("%03d", j))
      duplicated_row$rodne_uid <- paste0(duplicated_row$rodne_uid, "_", sprintf("%03d", j))
      
      # Set diam_cm to 7.5 for the duplicated row
      duplicated_row$diam_cm <- 7.5
      
      # Append the duplicated row to the list
      duplicated_rows[[length(duplicated_rows) + 1]] <- duplicated_row
    }
  }
  
  # If X5_10_cm has a value of 1, set diam_cm to 7.5
  if (!is.na(row$X5_10_cm) && row$X5_10_cm == 1) {
    duplicated_row <- row
    duplicated_row$diam_cm <- 7.5
    duplicated_rows[[length(duplicated_rows) + 1]] <- duplicated_row
  }
}

# Combine the duplicated rows into a dataframe
X5_10_cm_df <- do.call(rbind, duplicated_rows)

# Reset row names
rownames(X5_10_cm_df) <- NULL

################################################################################

# Bind the rows of the three data frames x0_5_cm_df, x5_10_cm_df, and tallied_df
tallied_all_df <- rbind(X0_5_cm_df, X5_10_cm_df, tallied_df)

# Reset row names
rownames(tallied_all_df) <- NULL

################################################################################

# Now we need to replace the id value rows within original_df with the ones that are in tallied_all_df

# Remove original rows with the same unique id values as tallied_all_df
original_df <- original_df[!original_df$id %in% tallied_all_df$id, ]

# Append all rows from tallied_all_df to original_df
original_df <- rbind(original_df, tallied_all_df)

# Reset row names
rownames(original_df) <- NULL
```

Correcting the naming schemes of canopy positions

```{r}
original_df <- original_df %>%
  group_by(pcid, trid) %>%
  mutate(capo_num = ifelse(is.na(capo_num), min(capo_num, na.rm = TRUE), capo_num)) %>%
  ungroup()

# Mapping the proper names
mapping <- c("Codominant" = "codominant",
             "CODOMINANT" = "codominant",
             "DEAD" = "dead",
             "Dominant" = "dominant",
             "DOMINANT" = "dominant",
             "Intermediate" = "intermediate",
             "INTERMEDIATE" = "intermediate",
             "Open Grown" = "dominant",
             "OPEN GROWN" = "dominant",
             "Suppressed" = "suppressed",
             "SUPPRESSED" = "suppressed",
             "Supressed" = "suppressed")

# Apply the mapping to the "capo" column
original_df <- original_df %>%
  mutate(capo = case_when(
    capo %in% names(mapping) ~ mapping[as.character(capo)],
    TRUE ~ as.character(capo)
  ))

# Checking the data
table(original_df$capo)

################################################################################

# Change the canopy descriptions so that they match the capo_num value
original_df <- original_df %>%
  mutate(capo = case_when(
    capo_num == 1 ~ "dominant",
    capo_num == 2 ~ "codominant",
    capo_num == 3 ~ "intermediate",
    capo_num == 4 ~ "suppressed",
    capo_num == 0 ~ "dead",
    TRUE ~ as.character(capo)  # Keep the original value if none of the other defined conditions match
  ))

# Checking the changes
table(original_df$capo)

################################################################################
# Change the value of capo_num to 4 where capo_num is 0 and cond is LIVE for data validaton
original_df$capo_num[original_df$capo_num == 0 & original_df$cond == "LIVE"] <- 4
```

Cleaning up column names and alternative tpi naming schemes

```{r}
# Clean up column names
original_df <- original_df %>%
  dplyr::rename(insol = Insolation,
         tpi = TPI,
         pvt = imap_name)

# Clean up the tpi rows
original_df$tpi[original_df$tpi == "MidSlope"] <- "Midslope"
original_df$tpi[original_df$tpi == "Midslope"] <- "Mid Slope"
original_df$tpi[original_df$tpi == "Ridge"] <- "Ridge Top"
table(original_df$tpi)
```

Refining the data and calculating diam_cm from rcdm (root crown diameter) for only the single stem trees

```{r}
# Get rid of the columns that are not needed (by keeping only the needed columns)
original_df <- original_df %>%
  select(id, site_name, rodne_site_ab, uid, rodne_uid, rodne_pid, plot_type, plot_size, 
         pcid, trid, stid, spcd, multistem, tally, diam_cm, dstmp, X5_10_cm, X0_5_cm, 
         no_diam, rcdm, indm, cond, corvallis_core, rodne_sid, outer_ring, inner_ring, pith, d2pith_mm,
         metlen_age, conifer, con_numeric, capo, capo_num, harvest_date, canopy_cov, tpi, x, y)

################################################################################
# Filter out the tree records that does not have any records of bole/root crown diameter
no_diam <- original_df %>%
  filter(is.na(diam_cm) & is.na(dstmp) & is.na(X5_10_cm) & is.na(X0_5_cm) & is.na(rcdm))

# Take the inverse of the no_diam dataframe for the final original_df that has some form of diameter measurement (RCDM)
original_df <- anti_join(original_df, no_diam, by = "rodne_uid")

################################################################################
# Now I can attempt to calculate diam_cm from rcdm for only the single stem trees
# Need to calculate the regression for diam_cm <- rcdm for single stemmed hardwood individuals
regr_hardwood <- original_df %>%
  filter(conifer=="Hardwood" & multistem==0 & tally==1)

# Generalized Additive Model for the regression equation to calculate diam_cm from rcdm
gammod1 <- gam(diam_cm ~ rcdm, data = regr_hardwood)
summary(gammod1)$r.sq
summary(gammod1)

# Create single_stemmed dataframe that will be calculated for diam_cm from rcdm
multi_stemmed <- original_df %>%
  filter(is.na(diam_cm) & rcdm>0) %>%
  mutate(diam_cm = ((rcdm*0.408434)+5.241285)/tally)

# Take the inverse of multi_stemmed to add back to original_df
non_multi <- anti_join(original_df, multi_stemmed, by = "rodne_uid")

# Join multi_stemmed and non_multi back into the original_df
original_df <- bind_rows(multi_stemmed, non_multi)

################################################################################
# Whenever tally is greater than 1, multistem is equal to 1
original_df <- original_df %>%
  mutate(multistem = ifelse(tally > 1, 1, multistem))
```

Converting distance to pith metric for cores from mm to cm

```{r}
################################################################################
# Convert the d2pith_mm to cm for the cored trees
original_df$d2pith_mm <- as.numeric(original_df$d2pith_mm)
original_df$d2pith_cm <- original_df$d2pith_mm / 10
```

Creating the rogue_trees file from live trees and the deathdated dead trees
Further refine the data columns and correct the data structure

```{r}
# Reduce the data frame columns to those needed for next steps
all_trees <- original_df[,c("site_name", "uid", "trid", 
                           "stid", "multistem", "rcdm", "conifer", "rodne_site_ab", "rodne_pid", "plot_type", "plot_size", "rodne_sid", "rodne_uid", "corvallis_core",
                           "spcd", "diam_cm", "capo_num", "cond", "con_numeric", "inner_ring", 
                           "outer_ring", "pith", "d2pith_cm", "metlen_age", "harvest_date", "tpi", "x", "y")]

# Correct the data structure
all_trees <- all_trees %>%
  mutate(.,
         rodne_pid = factor(rodne_pid),
         site_name = factor(site_name),
         uid = factor(uid),
         rodne_site_ab = factor(rodne_site_ab),
         plot_type = factor(plot_type),
         plot_size = factor(plot_size),
         rodne_sid = factor(rodne_sid),
         rodne_uid = factor(rodne_uid),
         corvallis_core = factor(corvallis_core),
         spcd = factor(spcd),
         diam_cm = as.numeric(diam_cm),
         capo_num = factor(capo_num),
         cond = factor(cond),
         con_numeric = factor(con_numeric),
         tpi = factor(tpi))

# All the placeholder pith values are called NA
all_trees$pith[all_trees$pith == -999] <- NA
```

Adding data to the main file that are needed for further analysis

```{r}
# add harvest year, plant association, pvt, bps, insolation values, and year sampled
invt_dates <- read.csv(here("csv_input_deck/plot_dates_plant_assoc_pvt_bps.csv"))
invt_dates <- invt_dates[c("rodne_pid", "field_plant_assoc", "field_pvt", "bps_name", "insol", "insol_value", "elevation_m", "year_sampled")]
table(invt_dates$field_pvt)

# Merge harvest years with the all trees data frame
all_trees <- merge(all_trees, invt_dates, by = "rodne_pid", all.x = TRUE)
```

Continuing to clean the data

```{r}
# Clean and shorten the column names
all_trees <- all_trees %>%
  dplyr::rename(plantassoc = field_plant_assoc,
         pvt = field_pvt,
         bps = bps_name,
         elevation = elevation_m)

# creating an age column
all_trees <- all_trees %>%
  mutate(true_age = ifelse(con_numeric == 1,
                           year_sampled - pith,
                           outer_ring - pith))

# Create the "undateable.csv" that contains all the trees that we cant date
non_aged_species <- c("ALRU", "CONU", "FRLA", "TSME", "UNKN")
undateable <- all_trees %>%
  filter(spcd %in% non_aged_species | is.na(diam_cm) | is.na(capo_num))

# Write the undateable csv (trees that are missing critical parameters) (n = 1678) These trees are unfortunately removed from further analysis
write.csv(undateable, 
          here("csv_output_deck/undateable.csv"),
          row.names=FALSE)

################################################################################
# Continue to create all_trees without the undateable
# all_trees contains all trees dead or alive with the filtered columns and not 
# including the unageable rows containing: ("ALRU", "CONU", "FRLA", "TSME", "UNKN") #5 species 
# or na in diam_cm and capo_num
age_species <- c("ABCO", "ARME", "CADE", "PIJE", "PILA", "PIPO", "PSME", "QUGA", "QUKE",
                 "ABMAS", "ACMA", "CHCH", "CHLA", "LIDE", "QUCH", "UMCA", "TABR")

all_trees <- all_trees %>%
  filter(spcd %in% age_species & !is.na(diam_cm) & !is.na(capo_num))

all_trees <- all_trees %>%
  mutate(spcd_mod = case_when(
    spcd %in% c("PIPO", "PIJE") ~ 1,
    spcd == "PILA" ~ 2,
    spcd %in% c("QUGA", "QUKE", "QUCH", "LIDE", "UMCA") ~ 3,
    spcd == "PSME" ~ 4,
    spcd %in% c("ABCO", "ABMAS") ~ 5,
    spcd %in% c("ARME", "ACMA") ~ 6,
    spcd %in% c("CADE", "CHLA", "CHCH", "TABR") ~ 7,
    TRUE ~ NA_integer_
  ))
```

Other climate variables like aet, pet, and water deficit are added to the main data set

```{r}
climate <- read.csv("csv_input_deck/rogue_climate_pvt.csv")

# Change the name for rodne_pid in climate dataframe
colnames(climate)[colnames(climate) == "Rodne_PID"] <- "rodne_pid"

# Remove duplicates
climate <- climate[!duplicated(climate$rodne_pid), ]

# Select specific columns from climate data frame
climate_selected <- climate[, c("rodne_pid", "WB_AET_Current_V2_2022", "WB_PET_Current_V2_2022", "WB_Deficit_Current_V2_2022_2023")]

# Join based on 'id'
all_trees <- merge(all_trees, climate_selected, by = "rodne_pid")

# Rename the climate variable columns
colnames(all_trees)[colnames(all_trees) == "WB_AET_Current_V2_2022"] <- "aet"
colnames(all_trees)[colnames(all_trees) == "WB_PET_Current_V2_2022"] <- "pet"
colnames(all_trees)[colnames(all_trees) == "WB_Deficit_Current_V2_2022_2023"] <- "deficit"

# Round the values
all_trees$aet <- round(all_trees$aet, 0)
all_trees$pet <- round(all_trees$pet, 0)
all_trees$deficit <- round(all_trees$deficit, 0)
```

Stand density is calculated and added as a possible analysis variable

```{r}
# Converting to inches
all_trees$diam_in <- (all_trees$diam_cm/2.54)

# Calculating the individual tree density values for the summation method of calculating stand density
all_trees$inddensity <- (all_trees$diam_in/10)^1.6

# Calculating basal area (sq ft) per acre in all_trees
all_trees <- all_trees %>%
  mutate(tree_area = (pi*(diam_in/2)^2)/144)

# Bringing in the stand density from "extra_estimating_stand_density.R"
density <- read.csv(here("csv_input_deck/stand_density_values.csv"))

# Select specific columns from density data frame
density_selected <- density[, c("rodne_pid", "tpa", "sdi_acre", "basal_area_acre")]

# Join based on 'id'
all_trees <- merge(all_trees, density_selected, by = "rodne_pid")

# Round off the sdi, basal area, and tpa values
all_trees$tpa <- round(all_trees$tpa, digits = 0)
all_trees$sdi_acre <- round(all_trees$sdi_acre, digits = 0)
all_trees$basal_area_acre <- round(all_trees$basal_area_acre, digits = 0)
```

Adding in the biostrat variable and still refining data for aging trees

```{r}
# Combining the biophysical characteristics into one column of stratum combos (biostrat)
all_trees$biostrat <- paste(all_trees$pvt, all_trees$insol, all_trees$tpi, sep = ",")
all_trees$biostrat = as.factor(all_trees$biostrat)

# Filtering out the unneeded columns
colnames(all_trees)
all_trees <- all_trees[, c("rodne_uid", "rodne_pid", "site_name",  "uid", "trid", 
                                    "stid", "multistem", "rcdm", "conifer",
                                      "rodne_site_ab", "plot_type",  "plot_size",  
                                      "rodne_sid",  "corvallis_core", "spcd", "diam_cm",  
                                      "capo_num", "cond", "con_numeric",  "inner_ring", 
                                      "outer_ring", "pith", "d2pith_cm", "metlen_age", "harvest_date", "plantassoc",
                                      "pvt", "bps", "insol", "insol_value", "tpi", "true_age", "year_sampled", 
                                      "spcd_mod", "aet", "pet", "deficit",  "elevation",  
                                      "diam_in",  "inddensity", "tree_area",  "tpa",  
                                      "sdi_acre", "basal_area_acre", "biostrat", "x", "y")]


# Write the "all_trees.csv" file that contains all trees
write.csv(all_trees, 
          here("csv_output_deck/all_trees.csv"),
          row.names=FALSE)

# Write "live_trees.csv" that contains all LIVE trees from the applicable species data set
live_trees <- all_trees %>%
  filter(spcd %in% age_species) %>%
  filter(cond=="LIVE")

write.csv(live_trees, 
          here("csv_output_deck/live_trees.csv"),
          row.names=FALSE)

################################################################################
# Write "dead_trees.csv" that contains all LIVE trees from the applicable species data set
dead_trees <- all_trees %>%
  filter(spcd %in% age_species) %>%
  filter(cond!="LIVE")
```

Calculating deathdates for trees

```{r}
# There are one of two ways to calculate the deathdates for the trees:
################################################################################
# Figure out when the trees died (Option 1: A mix of Rogers et al. 1984, Bakker 2008, Fule 1993, and Vandervlugt 2011)
# Option 2 is using snag fall rates calculated from the decaid data: Mellen et al. (2012) (this is the option used)

# Bringing in the live trees file again
live_trees <- read.csv("csv_output_deck/live_trees.csv")

# Calculating the snag fall rates
dated <- live_trees %>%
  filter(pith>1 & true_age>100)
snagfalldatedcount <- dated %>%
  dplyr::group_by(spcd_mod) %>% 
  dplyr::summarise(avg_dbh = mean(diam_cm))
snagfalldatedcount <- snagfalldatedcount %>%
  mutate(snaglife = 1 + (avg_dbh / 2))
snagfalldatedcount <- snagfalldatedcount %>%
  mutate(snagfall = 1 / (snaglife))

# Rename the snagfall rate column to snagfall_rogers
names(snagfalldatedcount)[names(snagfalldatedcount) == "snagfall"] <- "snagfall_rogers"

# Merge dead_trees with filtered snagfalldatedcount based on the species column
snagfalldatedcount <- snagfalldatedcount[, c("spcd_mod", "snagfall_rogers")]
dead_trees <- left_join(dead_trees, snagfalldatedcount, by = "spcd_mod")

# Continue to the snag ages part if using the Rogers method, if not continue calculating Mellen

################################################################################
# Another way to bring in snag fall rates from Option 2: Mellen et al. (2012)

# Bringing in the mellen_snag_rates.csv
snag_rates <- read.csv("csv_input_deck/mellen_snag_rates.csv")

# Editing the data
consolidated_snag_rates <- snag_rates %>%
  dplyr::group_by(species_group, slope_position, snag_dbh) %>%
  dplyr::summarise(avg_snag_fall_rate = mean(snag_fall_rate))

# Changing the column names to create duplicates in the dead_trees data frame
names(consolidated_snag_rates)[names(consolidated_snag_rates) == "species_group"] <- "snag_spcd_group"
names(consolidated_snag_rates)[names(consolidated_snag_rates) == "slope_position"] <- "snag_slope"
names(consolidated_snag_rates)[names(consolidated_snag_rates) == "avg_snag_fall_rate"] <- "snagfall_mellen"

# Now to give each dead tree a snag_fall_rate based on the match of data between
# the dead trees and the consolidated_snag_rates data frames
# Assigning the snag_spcd_group
dead_trees <- dead_trees %>%
  mutate(snag_spcd_group = case_when(
    spcd %in% c("PIPO", "PIJE") ~ "Ponderosa pine (OR)",
    spcd == "PILA" ~ "Western white pine",
    spcd %in% c("QUGA", "QUKE", "QUCH", "LIDE", "UMCA") ~ "Misc hardwoods",
    spcd == "PSME" ~ "Douglas-fir (east OR)",
    spcd %in% c("ABCO", "ABMAS") ~ "Grand/white fir",
    spcd == "ARME" ~ "Pacific madrone",
    spcd == "ACMA" ~ "Alder",
    spcd %in% c("CADE", "CHLA", "CHCH", "TABR") ~ "Western redcedar",
    TRUE ~ NA_character_
  ))

# Assigning the snag_slope
dead_trees <- dead_trees %>%
  mutate(snag_slope = case_when(
    tpi == "Ridge Top" ~ "Ridgetop",
    tpi %in% c("Bottom Slope", "Mid Slope") ~ "Other",
    TRUE ~ NA_character_
  ))

# Assigning the snag_dbh
dead_trees <- dead_trees %>%
  mutate(
    snag_dbh = ifelse(
      spcd %in% c("ARME", "PIJE", "PILA", "PIPO", "QUGA", "QUKE",
                  "ACMA", "LIDE", "QUCH", "UMCA", "TABR"),
      case_when(
        diam_cm < 20 ~ "Small",
        between(diam_cm, 20, 50) ~ "Medium",
        diam_cm > 50 ~ "Large",
        TRUE ~ NA_character_
      ),
      ifelse(
        spcd %in% c("ABCO", "CADE", "PSME", "ABMAS", "CHCH", "CHLA"),
        case_when(
          diam_cm < 25 ~ "Small",
          between(diam_cm, 25, 75) ~ "Medium",
          diam_cm > 75 ~ "Large",
          TRUE ~ NA_character_
        ),
        NA_character_)))

# Bring the consolidated_snag_rates snag_rates in to merge with dead_trees
# Left join the tree data frame with the consolidated data frame based on common columns
dead_trees <- merge(dead_trees, consolidated_snag_rates, by.x = c("snag_spcd_group", "snag_slope", "snag_dbh"), by.y = c("snag_spcd_group", "snag_slope", "snag_dbh"), all.x = TRUE)

################################################################################
# Now apply either the snag fall rate from the Rogers et al (the first data mutations here) or Mellen et al (second) methods/data
################################################################################
# 25%, 50%, and 75% Columns that determines how many years need to be taken from year_sampled
# For 25%
dead_trees <- dead_trees %>%
  mutate(
    decay_25_rogers = ifelse(con_numeric == 4,
                      ((log(.25) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 5,
                      ((log(.25) - log(1)) / log(1.15)) + 
                      ((log(.25) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 6,
                      ((log(.25) - log(1)) / log(1 + snagfall_rogers)) + 
                      ((log(.25) - log(1)) / log(1.15)) + 
                      ((log(.25) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 7,
                      ((log(.25) - log(1)) / log(1 + snagfall_rogers)) + 
                      ((log(.25) - log(1)) / log(1 + snagfall_rogers)) + 
                      ((log(.25) - log(1)) / log(1.15)) +
                      ((log(.25) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 8,
                      ((log(.25) - log(1)) / log(1 + snagfall_rogers)) +
                      ((log(.25) - log(1)) / log(1 + snagfall_rogers)) + 
                      ((log(.25) - log(1)) / log(1 + snagfall_rogers)) + 
                      ((log(.25) - log(1)) / log(1.15)) + 
                      ((log(.25) - log(1)) / log(1.2)),
                      NA_real_))))))  # Handle other cases (if any) with NA or specify a default value
    
# For 50%
dead_trees <- dead_trees %>%
  mutate(
    decay_50_rogers = ifelse(con_numeric == 4,
                      ((log(.50) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 5,
                      ((log(.50) - log(1)) / log(1.15)) + 
                      ((log(.50) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 6,
                      ((log(.50) - log(1)) / log(1 + snagfall_rogers)) + 
                      ((log(.50) - log(1)) / log(1.15)) + 
                      ((log(.50) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 7,
                      ((log(.50) - log(1)) / log(1 + snagfall_rogers)) + 
                      ((log(.50) - log(1)) / log(1 + snagfall_rogers)) + 
                      ((log(.50) - log(1)) / log(1.15)) +
                      ((log(.50) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 8,
                      ((log(.50) - log(1)) / log(1 + snagfall_rogers)) +
                      ((log(.50) - log(1)) / log(1 + snagfall_rogers)) + 
                      ((log(.50) - log(1)) / log(1 + snagfall_rogers)) + 
                      ((log(.50) - log(1)) / log(1.15)) + 
                      ((log(.50) - log(1)) / log(1.2)),
                      NA_real_))))))  # Handle other cases (if any) with NA or specify a default value

# For 75%
dead_trees <- dead_trees %>%
  mutate(
    decay_75_rogers = ifelse(con_numeric == 4,
                      ((log(.75) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 5,
                      ((log(.75) - log(1)) / log(1.15)) + 
                      ((log(.75) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 6,
                      ((log(.75) - log(1)) / log(1 + snagfall_rogers)) + 
                      ((log(.75) - log(1)) / log(1.15)) + 
                      ((log(.75) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 7,
                      ((log(.75) - log(1)) / log(1 + snagfall_rogers)) + 
                      ((log(.75) - log(1)) / log(1 + snagfall_rogers)) + 
                      ((log(.75) - log(1)) / log(1.15)) +
                      ((log(.75) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 8,
                      ((log(.75) - log(1)) / log(1 + snagfall_rogers)) +
                      ((log(.75) - log(1)) / log(1 + snagfall_rogers)) + 
                      ((log(.75) - log(1)) / log(1 + snagfall_rogers)) + 
                      ((log(.75) - log(1)) / log(1.15)) + 
                      ((log(.75) - log(1)) / log(1.2)),
                      NA_real_))))))  # Handle other cases (if any) with NA or specify a default value

################################################################################
# mellen et al
# 25%, 50%, and 75% Columns that determines how many years need to be taken from year_sampled
# For 25%
dead_trees <- dead_trees %>%
  mutate(
    decay_25_mellen = ifelse(con_numeric == 4,
                      ((log(.25) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 5,
                      ((log(.25) - log(1)) / log(1.15)) + 
                      ((log(.25) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 6,
                      ((log(.25) - log(1)) / log(1 + snagfall_mellen)) + 
                      ((log(.25) - log(1)) / log(1.15)) + 
                      ((log(.25) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 7,
                      ((log(.25) - log(1)) / log(1 + snagfall_mellen)) + 
                      ((log(.25) - log(1)) / log(1 + snagfall_mellen)) + 
                      ((log(.25) - log(1)) / log(1.15)) +
                      ((log(.25) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 8,
                      ((log(.25) - log(1)) / log(1 + snagfall_mellen)) +
                      ((log(.25) - log(1)) / log(1 + snagfall_mellen)) + 
                      ((log(.25) - log(1)) / log(1 + snagfall_mellen)) + 
                      ((log(.25) - log(1)) / log(1.15)) + 
                      ((log(.25) - log(1)) / log(1.2)),
                      NA_real_))))))  # Handle other cases (if any) with NA or specify a default value

# For 50%
dead_trees <- dead_trees %>%
  mutate(
    decay_50_mellen = ifelse(con_numeric == 4,
                      ((log(.50) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 5,
                      ((log(.50) - log(1)) / log(1.15)) + 
                      ((log(.50) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 6,
                      ((log(.50) - log(1)) / log(1 + snagfall_mellen)) + 
                      ((log(.50) - log(1)) / log(1.15)) + 
                      ((log(.50) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 7,
                      ((log(.50) - log(1)) / log(1 + snagfall_mellen)) + 
                      ((log(.50) - log(1)) / log(1 + snagfall_mellen)) + 
                      ((log(.50) - log(1)) / log(1.15)) +
                      ((log(.50) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 8,
                      ((log(.50) - log(1)) / log(1 + snagfall_mellen)) +
                      ((log(.50) - log(1)) / log(1 + snagfall_mellen)) + 
                      ((log(.50) - log(1)) / log(1 + snagfall_mellen)) + 
                      ((log(.50) - log(1)) / log(1.15)) + 
                      ((log(.50) - log(1)) / log(1.2)),
                      NA_real_))))))  # Handle other cases (if any) with NA or specify a default value

# For 75%
dead_trees <- dead_trees %>%
  mutate(
    decay_75_mellen = ifelse(con_numeric == 4,
                      ((log(.75) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 5,
                      ((log(.75) - log(1)) / log(1.15)) + 
                      ((log(.75) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 6,
                      ((log(.75) - log(1)) / log(1 + snagfall_mellen)) + 
                      ((log(.75) - log(1)) / log(1.15)) + 
                      ((log(.75) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 7,
                      ((log(.75) - log(1)) / log(1 + snagfall_mellen)) + 
                      ((log(.75) - log(1)) / log(1 + snagfall_mellen)) + 
                      ((log(.75) - log(1)) / log(1.15)) +
                      ((log(.75) - log(1)) / log(1.2)),
                      ifelse(con_numeric == 8,
                      ((log(.75) - log(1)) / log(1 + snagfall_mellen)) +
                      ((log(.75) - log(1)) / log(1 + snagfall_mellen)) + 
                      ((log(.75) - log(1)) / log(1 + snagfall_mellen)) + 
                      ((log(.75) - log(1)) / log(1.15)) + 
                      ((log(.75) - log(1)) / log(1.2)),
                      NA_real_))))))  # Handle other cases (if any) with NA or specify a default value

# Creating another column that is the endyear of each tree after applying the 50% year decay
dead_trees$year_sampled <- as.numeric(dead_trees$year_sampled)

# Rogers calculations of the 50% decomposition rate end year
dead_trees$endyear_rogers <- dead_trees$year_sampled + dead_trees$decay_50_rogers
dead_trees$endyear_rogers <- round(dead_trees$endyear_rogers)

# Mellen calculations of the 50% decomposition rate end year (preferred)
dead_trees$endyear_mellen <- dead_trees$year_sampled + dead_trees$decay_50_mellen
dead_trees$endyear_mellen <- round(dead_trees$endyear_mellen)

# Convert con_numeric variable to numeric
dead_trees <- dead_trees %>%
  mutate(con_numeric = as.numeric(as.character(con_numeric)))

# Generating the death years
dead_trees <- dead_trees %>%
  mutate(
    deathyear_rogers = case_when(
      con_numeric == 2 ~ harvest_date,
      con_numeric == 3 ~ year_sampled,
      con_numeric > 3 ~ endyear_rogers,
      TRUE ~ NA_real_
    ),
    deathyear_mellen = case_when(
      con_numeric == 2 ~ harvest_date,
      con_numeric == 3 ~ year_sampled,
      con_numeric > 3 ~ endyear_mellen,
      TRUE ~ NA_real_
    )
  )

################################################################################
# Write it!
write.csv(dead_trees, 
          here("csv_output_deck/dead_trees.csv"),
          row.names=FALSE)
```

Merging the calculated death dates with the main file

```{r}
################################################################################
# Correct data structure to bind rows of live and dead trees to rogue_trees for the KNN Algorithm
dead_trees <- dead_trees %>%
  mutate(.,
         rodne_pid = factor(rodne_pid),
         site_name = factor(site_name),
         uid = factor(uid),
         rodne_site_ab = factor(rodne_site_ab),
         plot_type = factor(plot_type),
         plot_size = factor(plot_size),
         rodne_sid = factor(rodne_sid),
         rodne_uid = factor(rodne_uid),
         corvallis_core = factor(corvallis_core),
         spcd = factor(spcd),
         diam_cm = as.numeric(diam_cm),
         capo_num = factor(capo_num),
         cond = factor(cond),
         con_numeric = factor(con_numeric),
         pvt = factor(pvt),
         insol = factor(insol),
         tpi = factor(tpi))

live_trees <- live_trees %>%
  mutate(.,
         rodne_pid = factor(rodne_pid),
         site_name = factor(site_name),
         uid = factor(uid),
         rodne_site_ab = factor(rodne_site_ab),
         plot_type = factor(plot_type),
         plot_size = factor(plot_size),
         rodne_sid = factor(rodne_sid),
         rodne_uid = factor(rodne_uid),
         corvallis_core = factor(corvallis_core),
         spcd = factor(spcd),
         diam_cm = as.numeric(diam_cm),
         capo_num = factor(capo_num),
         cond = factor(cond),
         con_numeric = factor(con_numeric),
         pvt = factor(pvt),
         insol = factor(insol),
         tpi = factor(tpi))

# Merging the live and dead trees into rogue trees
rogue_trees <- bind_rows(live_trees, dead_trees)

# Creating a new column that gives an end_date value for measurements (model_pith still serves as the theoretical start year)
rogue_trees$con_numeric <- as.numeric(rogue_trees$con_numeric)
rogue_trees$deathyear_mellen <- as.numeric(rogue_trees$deathyear_mellen)
rogue_trees$year_sampled <- as.numeric(rogue_trees$year_sampled)

# Making the end_year column (using the Mellen calulated death year)
rogue_trees <- rogue_trees %>%
  mutate(end_year = ifelse(con_numeric > 1, deathyear_mellen, ifelse(con_numeric == 1, year_sampled, NA)))
rogue_trees$end_year <- as.numeric(rogue_trees$end_year)
```

Writing "rogue_trees.csv" which is used as the main file for all subsequent modeling efforts

```{r}
# Creating the "rogue_trees.csv" that contains the aged live and dead trees
write.csv(rogue_trees, 
          "csv_output_deck/rogue_trees.csv",
          row.names=FALSE)
```

